Unique solutions of contractions, CCS, and their HOL formalisation\thanksref{ATL1}

### 2020-07-31 ic-journal/multivariate.tex
Updated generated LaTeX sources

The set of free and bound
variables of a term need not be disjoint, e.g.~in $X + \recu X E$.
More importantly, when going
from a CCS expression to its sub-expressions, the set of free
variables may increase, while the set of bound variables either
remains the same or decreases. For instance, $\fv{\recu X (\mu.X)} = \emptyset$, while
$\fv{\mu.X} = \{X\}$. This property of $\fv{\cdot}$ brings some
%% "transition induction" (1/2)
difficulities in transition inductions. As an evidence,
we prove the following fundamental property of
$\fv{\cdot}$~\citep[p.~1209]{milner1990operational}:
%
\begin{proposition}
\label{prop:transFV}
The derivatives of a process are themselves processes, i.e.
if $E \arr{\mu} E'$ and $\fv{E} = \emptyset$, then $\fv{E'} =
\emptyset$. Formally:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{IS_PROC} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{IS_PROC} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\hfill{[TRANS_PROC]}
\end{alltt}
\end{proposition}

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLTokenDefEquality{}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Variable substitution only occurs on free
variables. When the term $E$ of ``\HOLinline{\HOLConst{CCS_SUBST}\\\;\HOLFreeVar{fm}\\\;\HOLFreeVar{E}}''
is in the form of $\recu X E'$, if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E'$ using a reduced map~--- without $X$~--- so that all
occurences of $X$ in $E'$ are correctly bypassed.
Since \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted.

@@

\begin{lemma}[\texttt{CCS_SUBST_elim}]
If the free variables of $E$ is disjoint with \HOLinline{\HOLFreeVar{Xs}}, a
substitution of \HOLinline{\HOLFreeVar{Xs}} in $E$ does not change $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{E}
\end{alltt}
\end{lemma}

@@

It could be possible to inductively define \multivariate contexts and their
guarded variants, but care is needed for
variables that occur within a recusion. We have preferred a different
solution, in which the definitions of \multivariate contexts  is based on the
existing \univariate definitions.
For this, intuitively, we replace a single variable with a hole (viewing
the other variables as constants), so to obtain a  single-variable context
on which predicates like \HOLinline{\HOLConst{CONTEXT}} can be used.
For example, to see the weak guardedness of $a.X + b.X + c.Y$,
we substitute each occurrence of a variable with a ``hole'' and consider the
resulting term as a multi-hole context (i.e., a $\lambda$-function); then all such multi-hole contexts
should satisfy \HOLinline{\HOLConst{WG}}, i.e.~the following results hold:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}
\end{alltt}
Note that \HOLinline{\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\\\;\HOLSymConst{\ensuremath{=}}\\\;\ensuremath{[}\HOLFreeVar{t}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\ensuremath{(}\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\ensuremath{)}}
can be expressed as a \univariate substitution of the original term.
This idea leads to the following definitions:
\begin{alltt}
   \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLTokenDefEquality{} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}\hfill{[context_def]}

   \HOLConst{weakly_guarded} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLTokenDefEquality{} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}\hfill{[weakly_guarded_def]}
\end{alltt}
The above definitions take an extra list of variables \HOLinline{\HOLFreeVar{Xs}} and
assert the CCS expression \HOLinline{\HOLFreeVar{E}} with respect to this list. This
allows us to formalise the concepts of contexts and guardedness
independently of the free variables of \HOLinline{\HOLFreeVar{E}}.
Then a logical term ``\HOLinline{\HOLConst{weakly_guarded}\\\;\ensuremath{(}\HOLConst{SET_TO_LIST}\\\;\ensuremath{(}\HOLConst{FV}\\\;\HOLFreeVar{E}\ensuremath{)}\ensuremath{)}\\\;\HOLFreeVar{E}}'' can
be used to assert the weak guardedness of $E$ with respect to all its free
variables.\footnote{Here \HOLinline{\HOLConst{SET_TO_LIST}} converts a finite set
  to a list of the same elements. It turns out that we never need this, because
  in all unique-solution theorems a list of variables \HOLinline{\HOLFreeVar{Xs}} is fixed and
  then all equations are required to contain free variables in
  \HOLinline{\HOLFreeVar{Xs}}.}
(The \multivariate guardedness and sequentiality can also be defined similarily, using
their \univariate companions \HOLinline{\HOLConst{SG}} and \HOLinline{\HOLConst{SEQ}}.)

@@

With the formal definitions of \multivariate substitution and
\multivariate (weakly guarded) contexts, now we are ready to formally
define \multivariate CCS equations/contractions and their (unique)
solutions. Consider a system of equation $\{X_i = E_i\}_{i\in I}$
(Def.~\ref{def:equation}), or its expanded form (here we suppose $I = [1,n]
\in \mathbb{N}$, i.e.~a finite list):
\begin{equation*}
  \begin{cases}
    &X_1 = E_1[\til X] \\
    &X_2 = E_2[\til X] \\
    & \cdots \\
    &X_n = E_n[\til X]
  \end{cases}
\end{equation*}
Consider its two essential ingredients:
\begin{itemize}
\item $\til X = (X_1, X_2, \ldots, X_n)$: a list of equation variables;
\item $\til E = (E_1, E_2, \ldots, E_n)$: a list of CCS contexts
  with possible occurrences of free variables in $\til X$.
\end{itemize}
The two lists $\til X$ and $\til E$ should have the same
length, and  the variables in $\til X$ should  be distinct.
Furthermore, $\til E$ does not contain free variables that are not in $\til X$.
Finally, for each
$E_i$, the set of its bound variables should be \emph{disjoint} from $\til X$.
These requirements yield the fomal
definition of \multivariate CCS equation:
\begin{alltt}
   \HOLConst{CCS_equation} \HOLFreeVar{Xs} \HOLFreeVar{Es} \HOLTokenDefEquality{}
     \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
     \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{FV} \HOLBoundVar{e} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Es}
\end{alltt}

>>>>

The set of free and bound
variables of a term need not be disjoint, e.g.~in $X + \recu X E$.
More importantly, when going
from a CCS expression to its sub-expressions, the set of free
variables may increase, while the set of bound variables either
remains the same or decreases. For instance, $\fv{\recu X (\mu.X)} = \emptyset$, while
$\fv{\mu.X} = \{X\}$. This property of $\fv{\cdot}$ brings some
%% "transition induction" (1/2)
difficulties in transition inductions. As an evidence, % editorial changes
we prove the following fundamental property of
$\fv{\cdot}$~\citep[p.~1209]{milner1990operational}:
%
\begin{proposition}
\label{prop:transFV}
The derivatives of a process are themselves processes, i.e.
if $E \arr{\mu} E'$ and $\fv{E} = \emptyset$, then $\fv{E'} =
\emptyset$. Formally:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \ensuremath{\HOLBoundVar{E}\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{\HOLBoundVar{E}\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{IS_PROC} \HOLBoundVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLConst{IS_PROC} \ensuremath{\HOLBoundVar{E}\sp{\prime}}\hfill{[TRANS_PROC]}
\end{alltt}
\end{proposition}

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLTokenDefEquality{}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Variable substitution only occurs on free
variables. When the term $E$ of ``\HOLinline{\HOLConst{CCS_SUBST}\\\;\HOLFreeVar{fm}\\\;\HOLFreeVar{E}}''
is in the form of $\recu X E'$, if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E'$ using a reduced map~--- without $X$~--- so that all
occurrences of $X$ in $E'$ are correctly bypassed. % editorial changes
Since \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted.

@@

\begin{lemma}[\texttt{CCS_SUBST_elim}]
If the free variables of $E$ are % I&C editor issue 8
disjoint with \HOLinline{\HOLFreeVar{Xs}}, a
substitution of \HOLinline{\HOLFreeVar{Xs}} in $E$ does not change $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{E}
\end{alltt}
\end{lemma}

@@

It could be possible to inductively define \multivariate contexts and their
guarded variants, but care is needed for
variables that occur within a recursion. % I&C editor issue 9
We prefer % I&C editor issue 10
a different solution, in which the definitions of \multivariate contexts are % I&C editor issue 10
based on the existing \univariate definitions.
For this, intuitively, we replace a single variable with a hole (viewing
the other variables as constants), so to obtain a  single-variable context
on which predicates like \HOLinline{\HOLConst{CONTEXT}} can be used.
For example, to see the weak guardedness of $a.X + b.X + c.Y$,
we substitute each occurrence of a variable with a ``hole'' and consider the
resulting term as a multi-hole context (i.e., a $\lambda$-function); then all such multi-hole contexts
should satisfy \HOLinline{\HOLConst{WG}}, i.e.~the following results hold:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}
\end{alltt}
Note that \HOLinline{\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\\\;\HOLSymConst{\ensuremath{=}}\\\;\ensuremath{[}\HOLFreeVar{t}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\ensuremath{(}\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\ensuremath{)}}
can be expressed as a \univariate substitution of the original term.
This idea leads to the following definitions:
\begin{alltt}
   \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLTokenDefEquality{} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}\hfill{[context_def]}

   \HOLConst{weakly_guarded} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLTokenDefEquality{} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}\hfill{[weakly_guarded_def]}
\end{alltt}
The above definitions take an extra list of variables \HOLinline{\HOLFreeVar{Xs}} and
assert the CCS expression \HOLinline{\HOLFreeVar{E}} with respect to this list. This
allows us to formalise the concepts of contexts and guardedness
independently of the free variables of \HOLinline{\HOLFreeVar{E}}.
Then a logical term ``\HOLinline{\HOLConst{weakly_guarded}\\\;\ensuremath{(}\HOLConst{SET_TO_LIST}\\\;\ensuremath{(}\HOLConst{FV}\\\;\HOLFreeVar{E}\ensuremath{)}\ensuremath{)}\\\;\HOLFreeVar{E}}'' can
be used to assert the weak guardedness of $E$ with respect to all its free
variables.\footnote{Here \HOLinline{\HOLConst{SET_TO_LIST}} converts a finite set
  to a list of the same elements. It turns out that we never need this, because
  in all unique-solution theorems a list of variables \HOLinline{\HOLFreeVar{Xs}} is fixed and
  then all equations are required to contain free variables in
  \HOLinline{\HOLFreeVar{Xs}}.}
(The \multivariate guardedness and sequentiality can also be defined similarly, using % editorial
their \univariate companions \HOLinline{\HOLConst{SG}} and \HOLinline{\HOLConst{SEQ}}.)

@@

With the formal definitions of \multivariate substitution and
\multivariate (weakly guarded) contexts, now we are ready to formally
define \multivariate CCS equations/contractions and their (unique)
solutions. Consider a system of equation $\{X_i = E_i\}_{i\in I}$
(Def.~\ref{def:equation}), or its expanded form (here we suppose $I = [1,n]
\in \mathbb{N}$, i.e.~a finite list):
\begin{equation*}
  \begin{cases}
    &X_1 = E_1[\til X] \\
    &X_2 = E_2[\til X] \\
    & \cdots \\
    &X_n = E_n[\til X]
  \end{cases}
\end{equation*}
Consider its two essential ingredients:
\begin{itemize}
\item $\til X = (X_1, X_2, \ldots, X_n)$: a list of equation variables;
\item $\til E = (E_1, E_2, \ldots, E_n)$: a list of CCS contexts
  with possible occurrences of free variables in $\til X$.
\end{itemize}
The two lists $\til X$ and $\til E$ should have the same
length, and  the variables in $\til X$ should  be distinct.
Furthermore, $\til E$ does not contain free variables that are not in $\til X$.
Finally, for each
$E_i$, the set of its bound variables should be \emph{disjoint} from $\til X$.
These requirements yield the formal % editorial
definition of \multivariate CCS equation:
\begin{alltt}
   \HOLConst{CCS_equation} \HOLFreeVar{Xs} \HOLFreeVar{Es} \HOLTokenDefEquality{}
     \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
     \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{FV} \HOLBoundVar{e} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Es}
\end{alltt}
